

<!-- [![R-CMD-check](https://github.com/valeriehan6/scoreLR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/valeriehan6/scoreLR/actions/workflows/R-CMD-check.yaml) -->
<!-- [![Coverage status](https://codecov.io/gh/valeriehan6/scoreLR/branch/main/graph/badge.svg)](https://codecov.io/github/valeriehan6/scoreLR?branch=main) -->

# scoreLR

This package computes SLRs using various methods that address the dependence between the data and compares the various methods. 

## Installation

You can install the development version of scoreLR like so:

```{r}
remotes::install_github("https://github.com/valeriehan6/scoreLR")
```


## Example

How do you use this package?

Read a dataset in that follows these requirements:

```{r, eval=F}
exdata <- readr::read_csv("./data/shoe_data.csv")
```


Then create a train/test split that accounts for the dependence's using the dep_split function:
```{r}
exdata_split <- dep_split(exdata, 0.8, 20230413)
data <- exdata_split

KM_train <- data %>% filter(source1 == source2 & train == TRUE) # split based on indicator column
KM_test <- data %>% filter(source1 == source2 & train == FALSE) # split based on indicator column
# Filter data to rows where source1 != source2
KNM_train <- data %>% filter(source1 != source2 & train == TRUE)
KNM_test <- data %>% filter(source1 != source2 & train == FALSE)

```



Calculate SLRs based on various methods:
```{r}
ignore_dep(KM_train, KM_test, KNM_train, KNM_test)
avg_features(KM_train, KM_test, KNM_train, KNM_test)
```


You can do it all in one step with the slr_results function:
```{r}
slr_results(data)
```



## Source

Han, Valerie. "Classification performance of score-based likelihood ratios when data exhibit various degrees of dependence." (2022).
